---
title: "Exercise 14 - Fit a logistic regression model "
author: "Kolekar, Shilpa"
date: October 26th, 2020
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
## Set the working directory 
setwd("C:/Users/shilp/Documents/GitHub/dsc520/data")

## Load the `data/binary-classifier-data.csv` to
binary_classifer_df <- read.csv("binary-classifier-data.csv")
summary(binary_classifer_df)
```
## a. What is the accuracy of the logistic regression classifier?
```{r}
newModel <- glm(label ~ x + y, data = binary_classifer_df, family = binomial)

## install.packages("caTools")
library("caTools")

split <- sample(1:nrow(binary_classifer_df), 0.8 * nrow(binary_classifer_df))
train <- binary_classifer_df[split,]
test <- binary_classifer_df[-split,]

# run the test data through model
res <- predict(newModel, test, type= "response")
res <- predict(newModel, train, type = "response")

##validate the model confusion matrix
confmatrix <- table(Actual_Value = train$label, Predicted_Value = res > 0.5)
confmatrix

# Accuracy
(confmatrix[[1,1]]+ confmatrix[[2,2]]) / sum(confmatrix)

```
Accuracy if model is 58.10%.


## b. How does the accuracy of the logistic regression classifier compare to the nearest neighbors algorithm?
```{r}

##extract 1st column of train dataset because it will be used as 'cl' argument in knn function.
target_category <- binary_classifer_df[split,1]

##extract 1st column if test dataset to measure the accuracy
test_category <- binary_classifer_df[-split,1]

##load the package class
## install.packages("class")
library("class")

##run knn function, k=sqrt(1498)
test_pred <- knn(train,test,cl=target_category,k=39)
test_pred

##create confusion matrix
table <- table(test_category, test_pred)
table

# Accuracy
(table[[1,1]]+ table[[2,2]]) / sum(table)

```
Accuracy of the logistic regression classifier model is 58.10%, and the nearest neighbors algorithm model gave 97.00% of accuracy. 

## c. Why is the accuracy of the logistic regression classifier different from that of the nearest neighbors?
Nearest neighbors model has low biased and low variance data. To get the better accuracy rate on the model low biased (less error data in training data) and low variance (less error in the test data) is required. Based on the accuracy percentage, nearest neighbors model has low biased and low variance data.

